---
title: 'PSYCH 251: Replication Project Data Analysis Script'
author: "Neha Rajagopalan"
date: "2024-06-21"
output: html_document
---


## Data Loading, Cleaning and Filtering

### Loading Libraries

```{r}
library("tidyverse")
library("dplyr")
```

### Reading data of two pilot participants

```{r}
df.participant1 = read_csv("PilotA_Data/pilotA_data.csv") %>%
  slice(18:n())
df.participant2 = read_csv("PilotA_Data/pilotA_data1.csv") %>%
  slice(18:n())

df.participant = rbind(df.participant1, df.participant2)
df.participant
```

### Filtering data for relevant columns and values

```{r}
df.response = df.participant %>%
  select(rt, stimulus, response, task, correct_response, category) %>%
  filter(!is.na(correct_response)) %>%
  select(-c(stimulus))

df.response
```



### Retrieving correct answers and responses with appropriate response time

```{r}
# Here we filter for responses that have a response time between 100ms and 2000ms
df.correct = df.response %>%
  filter(as.numeric(rt) >= 100) %>%
  filter(as.numeric(rt) <= 2000) %>%
# Here we filter correct responses
  filter(response == correct_response) 

df.correct
```

Key for category codes:

Format: Same/Different(S/D) image | Mirror/Normal(M/N) orientation | French/Japanese/Faces/False/Tools (F/J/M/P/T) group

1. DMF - Different Mirror French
2. DMJ - Different Mirror Japanese
3. DMM - Different Mirror Faces
4. DMP - Different Mirror False
5. DMT - Different Mirror Tools
6. DNF - Different Normal French
7. DNJ - Different Normal Japanese
8. DNM - Different Normal Faces
9. DNP - Different Normal False
10. DNT - Different Normal Tools
11. SMF - Same Mirror French
12. SMJ - Same Mirror Japanese
13. SMM - Same Mirror Faces
14. SMP - Same Mirror False
15. SMT - Same Mirror Tools
16. SNF - Same Normal French
17. SNJ - Same Normal Japanese
18. SNM - Same Normal Faces
19. SNP - Same Normal False
20. SNT - Same Normal Tools


### Split category column into Group Categories (Japanese, French, Faces, False, Tools) and Same/Different - Mirror/Normal Categories

```{r}
df.correct$SDNM = substr(df.correct$category, 1, 2)
df.correct$group = substr(df.correct$category, 3, 3)

df.correct = df.correct %>%
  select(-c(category))
df.correct
```


## Data Analysis Script 

### Median correct response times summary and Category codes
```{r}
median_rt = df.correct %>%
  group_by(group, SDNM)%>%
  summarise(median_rt = median(as.numeric(rt), na.rm =FALSE))

median_rt
```
 


## ANOVA for median response times: 

```{r}
anova_results <- aov(median_rt$median_rt ~ median_rt$group)
summary(anova_results)
```

